Tree size, microhabitat, and hydraulic traits shape drought responses in a temperate broadleaf forest

### 2019-09-21 manuscript/climate_sensitivity_draft.Rmd
worked on discussion

Our analysis indicates that height is the primary factor through which tree size mediates drought response; however, specific identification of the primary physiological mechanism(s) is complicated by the fact that multiple drought-relevant environmental parameters and tree characteristics covary with tree height. To begin with, taller trees face inherent biophysical challenges in lifting water a greater distance against the effects of gravity and friction (**REFS**). They also face different microenvironments, which are partially related to canopy position. Even under non-drought conditions, evaporative demand increases with tree height (Fig. 1a-b; see also **REFS**. **The vegetation reaches higher temperatures[??](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/58)**, which would occur more during a drought when solar radiation tends to be higher and less water is available for evaporative cooling of the leaves. Furthermore, daytime $CO_2$ concentrations tend to be lower in the canopy than in the understory **(Fig. 2[??](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/58)** ([Koike et al. 2001](https://watermark.silverchair.com/21-12-13-951.pdf); MORE REFS), implying greater water costs of $CO_2$ uptake. **[McDowell-vertical profiles in isotopes?*]**. Taller trees are more likely to be in co-dominant or domiant canopy positions, although signficant decoupling can be introduced by the configuration of neighboring trees (Fig. 2**e?**) ([Mueller-Landau et al. 2016](https://www.ncbi.nlm.nih.gov/pubmed/16643303)). **[discuss $CP$] **

@@

3. height & canopy position - biophysical factors
  - and we show that height is more important than exposure, but doesn't eliminate the effect of it
  - We used crown position despite its uncertainty. However, height and crown position change relatively slowly so it shouldn't be that far off.
  - We did not use crowding index because it has more uncertainty
  - we see no indication that rooting volume gives a relative advantage to large trees. Root systems are important, but [Bretfeld et al](https://nph-onlinelibrary-wiley-com.smithsonian.idm.oclc.org/doi/full/10.1111/nph.15071) show that "the limiting physiological processes for transpiration shifted from operating at the plant–soil interface to the plant–atmosphere interface with progressing forest succession.""

>>>>

Our analysis indicates that height is the primary factor through which tree size mediates drought response; however, specific identification of the primary physiological mechanism(s) is complicated by the fact that multiple drought-relevant environmental parameters and tree characteristics covary with tree height. To begin with, taller trees face inherent biophysical challenges in lifting water a greater distance against the effects of gravity and friction (**REFS**). They also face different microenvironments, which are partially related to canopy position. Even under non-drought conditions, evaporative demand increases with tree height (Fig. 1a-b; see also **REFS**. **The vegetation reaches higher temperatures[??](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/58)**, which would occur more during a drought when solar radiation tends to be higher and less water is available for evaporative cooling of the leaves. Furthermore, daytime CO_2 concentrations tend to be lower in the canopy than in the understory **(Fig. 2[??](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/58)** ([Koike et al. 2001](https://watermark.silverchair.com/21-12-13-951.pdf); MORE REFS), implying greater water costs of $CO_2$ uptake. **[McDowell-vertical profiles in isotopes?*]**. Taller trees are more likely to be in domiant canopy positions, although signficant decoupling can be introduced by the configuration of neighboring trees (Fig. 2**e?**) ([Mueller-Landau et al. 2016](https://www.ncbi.nlm.nih.gov/pubmed/16643303)). We show that height is a far stronger predictor than crown position (Tables 1,4,5). Our analysis does have the limitation that canopy positions were recored in 2018 and undoubtedly changed for some trees since the 1960s, and we note that $CP$ became an increasingly poor predictor moving from 1999 back to 1966 (Table 4). However, rates of change are relatively slow, and the fact that some trees inevitably did grow towards more dominant positions would bias against the acceptance of (*H1.2*), implying that dominant crown positions did indeed reduce $R$, which makes sense in light of the vertical enviromnental gradients described above and agrees with previous studies showing greater drought sensitivity in more exposed trees ([@suarez_factors_2004]; [@scharnweber_confessions_2019]; MORE??).  It is safe to assume that currently suppressed trees have always been suppressed, and their relatively low $R$ (after correcting for height effects) is real, which is consistent with analyses showing that suppressed--and particularly crowded--trees can suffer distproportionately during drought (**REFS**). The observed height-sensitivity of $R$, together with the apparent lack of importance of root water access (*H1.3*), agrees with the concept that physiological limitations to transpiration under drought shift from root water access to the plant-atmosphere intreface as forests age ([Bretfeld et al](https://nph-onlinelibrary-wiley-com.smithsonian.idm.oclc.org/doi/full/10.1111/nph.15071)), such that tall trees--particularly the very tallest--are the most sensitive in mature forests.

@@

3. height & canopy position - biophysical factors
